\section{Arbitration Brief Document Class}
 The file _arbitrationbrief.cls_ provides a document class (derived from _memoir_) for arbitration-style briefs, such as is used
 in the annual Willem C. Vis International Arbitration Moot. It does not actually share much in the way
 of code with other LawTeX packages, but is included due to it's related subject matter.

 This class file was used for the Yale Law School respondant's submission to the 2012 Vis Moot. Much of the boilerplate text 
 (title page, headers, etc.) do not have commands provided to adapt them to other uses, therefore you will need to edit these 
 portions of the class file yourself.


\cmd{newauthority}{Type}{Short Name}{Full Name}[Declare a new authority]

 \noindent This is the _arbitrationbrief.cls_ equivalent of the LawTeX _\newcase_, etc. New authorities are declared with this command, to which one
 must provide an Authority Type (Commentary, Statutes, Rules, or Cases), a short name to appear in the body of the text, and a long name to 
 appear in the table of authorities.


\cmd{cite}{Short Name}[Cite an authority in a citation block]

 \noindent This adds a citation to the text, ensuring that it will be recorded in the table of authorities, and providing a hyperlink thereto.
 This command does not take a pin cite; that should be given explicitly in the flow of the text, if appropriate.


\cmd{romancite}{Short Name}[Cite an authority in the body text.]

 \noindent This adds a citation to the text, in roman (not italic) font, ensuring that it will be recorded in the table of authorities, and providing a hyperlink thereto.
 This command does not take a pin cite; that should be given explicitly in the flow of the text, if appropriate. 

