\section{Law Brief Document Class}
The file lawbrief.cls provides a flexible document class for appellate-style legal briefs,
including cover page, table of contents, tables of authorities, etc.
 \subsection{Sectioning}
 We redefine formatting for section commands and the associated numbering. By default,
 the top-level _\section_ (Question Presented, Statement of Facts, Argument, etc.) is unnumbered, bold and centered;
 _\subsection_ is bold and left-aligned with a roman numeral; _\subsubsection_ is bold, left-indented, and with a letter
 for ordering. (Note that the formatting in the table of contents differs somewhat, see below). By default, 
 _\hyphenpenalty=10000_ is set so as to disable any hyphenation of words in each heading.

 \subsection{Configuring the Index / Table of Authorities}
 

_\indexindentsize_ --- Hanging indentation of toa from the left side (default: \the\indexindentsize)


\noindent _\indexrightmargin_ --- Right margin for all-but-the-last lines of toa (default: \the\indexrightmargin)
 The current value will preserve .5 inch of space for non-final lines, and ragged margins. To change the spacing on
 the last line between the entry and the page numbers, you must edit the style file.

\noindent _\passimlimit_ --- Use ``passim'' above this many pages (default: \passimlimit).  
 Can be changed with \\ _\renewcommand{\passimlimit}{5}_, etc. 
 Make sure to set this before the index is created (i.e. in the header).


\noindent _\idxpassim_ --- Used in an index command, forces passim to always be used for a particular cite

\noindent
_\caseindextitle_ --- Set the title for the case section of the table of authorities (Default: \@caseindextitle).

\noindent
_\statutesindextitle_ --- Set the title for the statutes section of the table of authorities (Default: \@statutesindextitle).

\noindent
_\otherindextitle_ --- Set the title for the miscellaneous section of the table of authorities (Default: \@otherindextitle).


 \noindent _\tableofauthorities_ --- Print the table of authorities.


 \subsection{Table-of-Contents commands}


_\tableofcontents_ --- Print the table of contents.


\noindent _\@tocline_ --- Sets the leaders to print aligned dots. Redefine to change.


\noindent _\l@section_ --- Formats Top-level headings in the ToC flush left and small caps


\noindent _\l@subsection_ --- Formats Second-level headings indented from left with a hanging indent of _\@pnumwidth_


\noindent _\l@subsubsection_ --- Formats Third-level headings indented more from left with a hanging indent of _\@pnumwidth_


 \subsection{Page numbering commands}

 These commands are used by calling _\pagestyle{romanparen}_, _\thispagestyle{toa}_, etc.


 \noindent Pagestyle _arabicparen_ --- arabic numerals inside parenthesis at bottom


 \noindent Pagestyle _romanparen_ --- roman numerals inside parenthesis at bottom


 \noindent Pagestyle _toa_ --- ``inherits'' from romanparen, but also puts the current mark at the top


 \noindent Pagestyle _footertext_ --- Sets the argument to _\footertext_ in the footer, flush left

 \subsection{Title Page commands}

 \noindent _\firstparty_, _\secondparty_ --- Sets the name of the first and second parties in the title page or caption.

 \noindent _\firstpartytitle_, _\secondpartytitle_ --- Set the title (Plaintiff, Defendants, etc.) of the parties.

 \noindent _\plaintiff_, _\defendant_, _\appellant_, _\appellee_, _\petitioner_, _\respondant_ --- Set both the name and title of each party. 
 Note that each of these also has a plural version (e.g., _\plaintiffs_).

 \noindent _\titlegraphic_ --- Set the stylized header for the title page.

 \noindent _\maketitle_ 

 This creates the title page, and generally would be called immediately
 after the _\begin{document}_. Alternatively, you may call _\makefrontmatter_, which 
 will create the title, question presented (which per Supreme Court rules is on the page 
 immediately after the title), table of contents, and table of authorities.

 For additional changes, such as an overall formatting change, I suggest that 
 it would be easiest for you to modify the macro _\maketitle_ or _\makefrontmatter_ 
 in the class file directly.

 \noindent _\makecaption_ --- Creates a trial-brief style caption, as opposed to a cover page.


 \noindent _\makefrontmatter_

 Creates all the boilerplate front matter, including the title page (by calling _\maketitle_), 
 the question presented (taken from the argument of _\questionpresented_), the table of contents and 
 the table of authorities (by calling _\tableofcontents_ and _\tableofauthorities_, respectively).


 \noindent _\rightbox_ environment

 The contents of this environment will themselves be left-aligned, but the entire box 
 containing them will be as far to the right as possible. This is useful for signature-type
 boxes at the end of the brief.

